1.6: Optimize Workflow: Projects, Markdown, and Git

### 2019-09-13 content/r/05-r-practice.Rmd
Test fixing footnote in r practice 5
Update slides 6, class 6, reading 7, add problem set 1 files, r practice 5 answers

@@

```{r, eval=F}
ifelse(my_conditions,
       yes = do_this_if_TRUE,
       no = do_this_if_FALSE)
```
]

>>>>

- [<i class="fas fa-external-link-square-alt"></i> `Answer Key (html)`](/r/05-r-practice-answers.html)
- [<i class="fab fa-markdown"></i> `Answer Key (markdown)`](/r/05-r-practice-answers.Rmd)
- [<i class="fas fa-registered"></i> `Answer Key (R Script)`](/r/05-r-practice-answers.R)

@@

`ifelse(my_conditions,
       yes = do_this_if_TRUE,
       no = do_this_if_FALSE)`]

### 2019-09-13 content/reading/07-reading.Rmd
Update slides 6, class 6, reading 7, add problem set 1 files, r practice 5 answers

- <i class="fas fa-external-link-square-alt"></i>[Chs. 8, 27, 28, 29, 30 (but especially 27)](http://r4ds.had.co.nz) in @R4DS

# Other Resources

## On R Markdown
- Xie, Yihui, J.J. Allaire, and Garett Grolemund, [*R Markdown: The Definitive Guide*](https://bookdown.org/yihui/rmarkdown/)
- RStudio, [*R Markdown Reference Guide*](https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf)
- RStudio, [*Overview of R Markdown*](https://rmarkdown.rstudio.com/)
- Rstudio, [*R Markdown Cheatsheet](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf)

## On Optimized Workflow

- Healey, Kieran, [*The Plain Person’s Guide to Plain Text Social Science*](http://plain-text.co/)
- Bryan, Jenny and Jim Hester, [*Happy Git and GitHub for the useR*](https://happygitwithr.com/)

>>>>

- <i class="fas fa-book"></i> @Bailey[Math and Probability Background Appendices A-I]

Now that we return to the statistics, we will do a minimal overview of basic statistics and distributions. Review Bailey's appendices.

### 2019-09-13 static/homeworks/01-problem-set-pdf.Rmd
Update slides 6, class 6, reading 7, add problem set 1 files, r practice 5 answers

---
title: "Problem Set 1"
editor_options: 
  chunk_output_type: console
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(pander)
source(file.path(here::here(), "lib", "graphics.R"))
```

*Due by Thursday, September 19, 2019*

- [<i class="fas fa-file-pdf"></i> `Download as PDF`](/homeworks/01-problem-set.pdf)
- [<i class="fab fa-markdown"></i> `Download Markdown`](/homeworks/01-r-problem-set.Rmd)

@@

## 1. 
### a. 
What are the top 5 boys names for 2017, and what *percent* of overall names is each?

### b.
What are the top 5 *girls* names, and what *percent* of overall names is each?

## 2. 
Make two barplots, of these top 5 names, one for each sex. Map `aes`thetics `x` to `name` and `y` to `prop`^[Or `percent`, if you made that variable, as I did.] and use `geom_col` (since you are declaring a specific `y`, otherwise you could just use `geom_bar()` and just an `x`.)

## 3. 
Find your name.^[If your name isn't in there :(, pick a random name.] `count` by `sex` how many babies since 1880 were named your name.^[Hint: if you do this, you'll get the number of *rows* (years) there are in the data. You want to add the number of babies in each row (`n`), so inside `count`, add `wt=n` to weight the count by `n`.] Also add a variable for the percent of each sex.

## 4.
Make a line graph of the number of babies with your name over time, `color`ed by `sex`.

## 5.
### a.
Make a table of the most common name for boys by year between 1980-2017.^[Hint: once you've got all the right conditions, you'll get a table with a lot of data. You only want to `slice` the `1`st row for each table.]

### b.
Now do the same for girls.

## 6. Now let's graph the evolution of the most common names since 1880.
### a.
First, find out what are the top 10 *overall* most popular names for boys and for girls. You may want to create two vectors, each with these top 5 names.

### b.
Now make two `line`graphs of these 5 names over time, one for boys, and one for girls.

## 7.
**Bonus (hard!)**: What are the 10 most common "gender-neutral" names?^[This is hard to define. For our purposes, let's define this as names where between 48 and 52% of the babies with the name are Male.]

@@

## 8. 
Download these two datasets that I've cleaned up a bit:^[If you want, try downloading them from the websites yourself!]

@@

Load them with `df<-read_csv("name_of_the_file.csv")` and save one as `econfreedom` and the other as `polfreedom`. Look at each `tibble` you've created.

## 9. 
The `polfreedom` dataset is still a bit messy. Let's overwrite it (or assign to something like `polfreedom2`) and select `Country.Territory` and `Total` (total freedom score) and rename `Country.Territory` to `Country`.

## 10.
Now we can try to merge these two datasets into one. Since they both have `Country` as a variable, we can merge these tibbles using `left_join(econfreedom, polfreedom, by="Country")`^[Note, if you saved as something else in question 9., use that instead of `polfreedom`!] and save this as a new tibble (something like `freedom`).

## 11.
Now make a scatterplot of Political Freedom (`total`)^[Feel free to `rename` these!] as `y` on Economic Freedom (`ef`) as `x` and `color` by `continent`.

## 12.
Let's do this again, but highlight some key countries. Pick three countries, and make a new tibble from `freedom` that is only the observations of those countries. Additionally, *install* and *load* a packaged called `ggrepel`^[This automatically adjusts labels so they don't cover points on a plot!] Next, redo your plot from question 11, but now add a layer: `geom_label_repel` and set its `data` to your three-country tibble, use same `aes`thetics as your overall plot, but be sure to add `label = ISO`, to use the ISO country code to label.^[You might also want to set a low `alpha` level to make sure the labels don't obscure other points!]

## 13.
Make another plot similar to 12, except this time use GDP per Capita (`gdp`) as `y`. Feel free to try to put a regression line with `geom_smooth()`!^[If you do, be sure to set its data to the full `freedom`, not just your three countries!]^[Those of you in my Development course, you just made my graphs from [Lesson 2](https://devf19.classes.ryansafner.com/slides/02-slides#23)!]

>>>>

---
title: "Problem Set 1"
author: "Ryan Safner"
date: "ECON 480 - Fall 2019"
output: pdf_document
---

@@

1. 
a. What are the top 5 boys names for 2017, and what *percent* of overall names is each?

\vspace{3in}

b. What are the top 5 *girls* names, and what *percent* of overall names is each?

\vspace{3in}

2. Make two barplots, of these top 5 names, one for each sex. Map `aes`thetics `x` to `name` and `y` to `prop`^[Or `percent`, if you made that variable, as I did.] and use `geom_col` (since you are declaring a specific `y`, otherwise you could just use `geom_bar()` and just an `x`.)

\clearpage

3. Find your name.^[If your name isn't in there :(, pick a random name.] `count` by `sex` how many babies since 1880 were named your name.^[Hint: if you do this, you'll get the number of *rows* (years) there are in the data. You want to add the number of babies in each row (`n`), so inside `count`, add `wt=n` to weight the count by `n`.] Also add a variable for the percent of each sex.

\vspace{3in}

4. Make a line graph of the number of babies with your name over time, `color`ed by `sex`.

\clearpage
5.
a. Make a table of the most common name for boys by year between 1980-2017.^[Hint: once you've got all the right conditions, you'll get a table with a lot of data. You only want to `slice` the `1`st row for each table.]

\vspace{3in}

b. Now do the same for girls.

\clearpage

6. Now let's graph the evolution of the most common names since 1880.
a. First, find out what are the top 10 *overall* most popular names for boys and for girls. You may want to create two vectors, each with these top 5 names.

b. Now make two `line`graphs of these 5 names over time, one for boys, and one for girls.

\vspace{3in}

7. **Bonus (hard!)**: What are the 10 most common "gender-neutral" names?^[This is hard to define. For our purposes, let's define this as names where between 48 and 52% of the babies with the name are Male.]

\clearpage

@@

8. Download these two datasets that I've cleaned up a bit:^[If you want, try downloading them from the websites yourself!]

@@

9. Load them with `df<-read_csv("name_of_the_file.csv")` and save one as `econfreedom` and the other as `polfreedom`. Look at each `tibble` you've created.

10.  The `polfreedom` dataset is still a bit messy. Let's overwrite it (or assign to something like `polfreedom2`) and select `Country.Territory` and `Total` (total freedom score) and rename `Country.Territory` to `Country`.

11. Now we can try to merge these two datasets into one. Since they both have `Country` as a variable, we can merge these tibbles using `left_join(econfreedom, polfreedom, by="Country")`^[Note, if you saved as something else in question 9., use that instead of `polfreedom`!] and save this as a new tibble (something like `freedom`).

\vspace{3in}

12. Now make a scatterplot of Political Freedom (`total`)^[Feel free to `rename` these!] as `y` on Economic Freedom (`ef`) as `x` and `color` by `continent`.

\clearpage

13. Let's do this again, but highlight some key countries. Pick three countries, and make a new tibble from `freedom` that is only the observations of those countries. Additionally, *install* and *load* a packaged called `ggrepel`^[This automatically adjusts labels so they don't cover points on a plot!] Next, redo your plot from question 11, but now add a layer: `geom_label_repel` and set its `data` to your three-country tibble, use same `aes`thetics as your overall plot, but be sure to add `label = ISO`, to use the ISO country code to label.^[You might also want to set a low `alpha` level to make sure the labels don't obscure other points!]

\vspace{2in}

14.  Make another plot similar to f, except this time use GDP per Capita (`gdp`) as `y`. Feel free to try to put a regression line with `geom_smooth()`!^[If you do, be sure to set its data to the full `freedom`, not just your three countries!] Those of you in my Development course, you just made my graphs from [Lesson 2](https://devf19.classes.ryansafner.com/slides/02-slides#23)!

### 2019-09-13 static/slides/06-slides.Rmd
Update slides 6, class 6, reading 7, add problem set 1 files, r practice 5 answers

1. Create a new repository on Github.
2. Start a New R Project in R Studio (link it to the github repository - [see guide](http://happygitwithr.com))
3. Create a logical file system ([see example](https://github.com/ryansafner/workflow)), such as:
```
project # folder on my computer (the new working directory)
|
|- Data/ # folder for data files 
|- Scripts/ # folder .R code
|- Bibliography/ # folder for .bib files
|- Figures/ # folder to plots and figures to
|- paper.Rmd # write document here
```
4. Write document in `paper.Rmd`, loading/saving files from/to various folders in project
5. Knit document to `pdf` or `html`.
6. Occasionally, `stage` and `commit` changes with a description, `push` to GitHub.

>>>>

1. Create a new repository on Github.<sup>.red[*]</sup>
2. Start a New R Project in R Studio (link it to the github repository<sup>.red[*]</sup> - [see guide](http://happygitwithr.com))
3. Create a logical file system ([see example](https://github.com/ryansafner/workflow)), such as:
```
project # folder on my computer (the new working directory)
|
|- Data/ # folder for data files 
|- Scripts/ # folder .R code
|- Bibliography/ # folder for .bib files
|- Figures/ # folder to plots and figures to
|- paper.Rmd # write document here
```
4. Write document in `paper.Rmd`, loading/saving files from/to various folders in project
    - e.g. load data like `df<-read_csv("Data/my_data")`; save plots like `ggsave("Figures/p.png")`
5. Knit document to `pdf` or `html`.
6. Occasionally, `stage` and `commit` changes with a description, `push` to GitHub.<sup>.red[*]</sup>

.footnote[<sup>.red[*]</sup> Optional and a bit advanced, remember this is _my_ workflow.]
---
